---
title: "Non-inferiority Goat"
date: " `r format(Sys.time(), '%d %B, %Y')` "
output:
  word_document:
    toc: true
    toc_depth: 2
    fig_caption: true
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_caption: true
    number_sections: true
  pdf_document:
    toc: true
    toc_depth: 2
    fig_caption: true
    number_sections: true
fontsize: 12pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

packages<-c("readxl", "nlme", "lme4","lmerTest","car","lattice","insight","performance",
            "HLMdiag","ggplot2", "sjPlot","sjmisc","emmeans","multcomp","multcompView",
            "lmtest", "rcompanion","MASS","blmeco", "knitr", "stargazer", "dplyr", "plyr",
            "summarytools", "table1", "survival","survminer", "glmmTMB", "brglm2")

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

# Load packages
ipak(packages)
```


```{r, include=FALSE}
datahalf = read_excel("1Final Master Efficacy 8.26.24.xlsx", 
                  sheet = "Half-level data", na = ".") 

datagoat = read_excel("1Final Master Efficacy 8.26.24.xlsx", 
                  sheet = "Goat-level data values", na = ".") 

datapredry = read_excel("1Final Master Efficacy 8.26.24.xlsx", 
                  sheet = "All DO cultures combined", na = ".") 

#str(datahalf)
#str(datagoat)
```

```{r, include=FALSE}
datahalf$DoeID <- as.factor(datahalf$DoeID)
datahalf$Farm <- as.factor(datahalf$Farm)
datahalf$Treatment <- as.factor(datahalf$Treatment)
datahalf$Treatment <- relevel(datahalf$Treatment, ref = "Control")
datahalf$Half <- as.factor(datahalf$Half)

datahalf$IMIatDO <- as.factor(datahalf$IMIatDO)
datahalf$IMIatPF <- as.factor(datahalf$IMIatPF)
datahalf$Cure <- as.factor(datahalf$Cure)
datahalf$NewIMI <- as.factor(datahalf$NewIMI)

#str(datahalf)

datagoat$DoeID  <- as.factor(datagoat$DoeID)
datagoat$Farm  <- as.factor(datagoat$Farm )
datagoat$Treatment  <- as.factor(datagoat$Treatment)
datagoat$Treatment <- relevel(datagoat$Treatment, ref = "Control")
datagoat$Parity <- as.factor(datagoat$Parity)

datagoat$MastitisEvent  <- as.factor(datagoat$MastitisEvent )
datagoat$Death  <- as.factor(datagoat$Death)
datagoat$Culling  <- as.factor(datagoat$Culling )
datagoat$DeathCulling  <- as.factor(datagoat$DeathCulling )

datagoat$SurvMastitis  <- as.numeric(datagoat$MastitisEvent)
datagoat$SurvDeath  <- as.numeric(datagoat$Death)
datagoat$SurvCulling  <- as.numeric(datagoat$Culling)
datagoat$SurvDeathCulling  <- as.numeric(datagoat$DeathCulling )


datapredry$Farm <- as.factor(datapredry$Farm)
datapredry$CultureResult <- as.factor(datapredry$CultureResult)
#str(datagoat)

```

# Data analysis - Quarter-level dynamics:

```{r, include=FALSE}
Desc1 <- datahalf %>% subset(select=c( Farm,	DoeID,	Treatment,	Half,	IMIatDO,	IMIatPF,	Cure,	NewIMI )) 
```

## Descriptive statistics - Culture Pre-dry:

### Overall:
```{r, echo=FALSE}

table1::table1(~ CultureResult , data=datapredry)

```

### By Farm:
```{r, echo=FALSE}

table1::table1(~ CultureResult | Farm, data=datapredry)

```


## Descriptive statistics - Culture DO:

### By Farm:

```{r, echo=FALSE}

table1::table1(~ DOResult | Farm, data=datahalf)

```

### By Treatment:

```{r, echo=FALSE}

table1::table1(~ DOResult | Treatment, data=datahalf)

```

### Treatment By Farm:

```{r, echo=FALSE}

table1::table1(~ DOResult | Treatment*Farm, data=datahalf)

```


## Descriptive statistics - Culture PF:

### By Farm:

```{r, echo=FALSE}

table1::table1(~ PFResult	| Farm, data=datahalf)
```

### By Treatment:

```{r, echo=FALSE}

table1::table1(~ PFResult	| Treatment, data=datahalf)
```

### Treatment By Farm: 

```{r, echo=FALSE}

table1::table1(~ PFResult	| Treatment*Farm, data=datahalf)
```


## Model 1 - IMI at DO:

Statistical model: generalized linear model (logistic regression)
  
  - Outcome: IMI at DO.
  - Fixed effect: Treatment.
  - Covariate: No covariates available.
  - Random effects: None.
    - Doe within farms were removed from random because little variation was explained adding it.

**Model:**

```{r , echo=FALSE}
#model1 <- glmmTMB(as.numeric(IMIatDO) ~ Treatment + (1|Farm) , 
#           family = poisson,
#           data = datahalf)
library(logistf)

#model1 <- glm(IMIatDO ~ Treatment,  #+ (1| Farm), 
#           family = binomial(link="logit"), 
#           #control = glmerControl(optimizer = "bobyqa"),
#           data = datahalf)
#summary(model1)
#print(model1, digits=3, corr=FALSE)
#tab_model(model1)


model1 <- logistf(IMIatDO ~ Treatment, data=datahalf)
summary(model1)

#model2 <- logistf(IMIatDO ~ Farm + Treatment, data=datahalf)
#summary(model2)



```


**LS Means - Prabability (%):**

```{r , echo=FALSE}

marginal.bin = lsmeans(model1, ~ Treatment)

emmeans(model1, "Treatment", type = "response")

b <- cld(marginal.bin, alpha   = 0.05, Letters = letters,    
    type    = "response", ### Report emmeans in orginal scale
    #adjust =  "sidak",
    decreasing = TRUE)

options(digits=3)

kable(b)
```


**LS Means - Prabability (%):**

```{r , echo=FALSE}

marginal.bin = lsmeans(model2, ~ Farm)

emmeans(model2, "Farm", type = "response")

b <- cld(marginal.bin, alpha   = 0.05, Letters = letters,    
    type    = "response", ### Report emmeans in orginal scale
    #adjust =  "sidak",
    decreasing = TRUE)

options(digits=3)

kable(b)
```


**Differences between treatments:**

```{r , echo=FALSE}

betta = as.data.frame(pairs(marginal.bin, adjust="tukey"))

kable(marginal.bin)
kable(betta)

```


**LS Means - Plot:**

```{r, echo=FALSE}
ggplot(b,
       aes(x     = Treatment,
           y     = prob,
           color = Treatment,
           label = .group, 
           ymax=1, ymin=0)) +
    geom_point(shape  = 15, size   = 4) +
    geom_errorbar(aes(ymin  =  lower.CL	,
                      ymax  =  upper.CL),
                      width =  0.2,
                      size  =  0.7) +
     theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
    ylab("Probability, %")+
    xlab("Treatment Group")+
   scale_color_manual(values = c("#999999", "#000000","blue", "red" ))+
  theme_bw() +
  theme(legend.position="bottom")


```


**Estimate of risk and risk difference against the NT:**

```{r}
options(digits=5)

### Odds ratio difference: NT - CLOX (Orbenin)
table(datahalf$Treatment, datahalf$IMIatDO)

pr1 = 1.0 
n1 = 129 #Reference
pr2 = 1.0 
n2 = 136

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

(azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100))



### Odds ratio difference: NT - CEPH (Tomorrow)
table(datahalf$Treatment, datahalf$IMIatDO)

pr1 = 1.0  
n1 = 129 #Reference
pr2 = 1.0 
n2 = 138

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 


### Odds ratio difference: NT - NC
table(datahalf$Treatment, datahalf$IMIatDO)

pr1 = 1.0  
n1 = 129 #Reference
pr2 = 0.003
n2 = 175

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 
```


## Model 2 - IMI at PK:

Statistical model: generalized linear mixed model (logistic regression)

  - Outcome: IMIatPF.
  - Fixed effect: Treatment.
  - Covariate: IMI at D0 tested, but removed since was not significant.
  - Random effects: Doe within Farms.

**Regression**:

```{r , echo=FALSE}

model2 <- glmer(IMIatPF ~ Treatment   + (1| Farm/DoeID), 
           family = binomial(link="logit"), 
           control = glmerControl(optimizer = "bobyqa"),
           data = datahalf)

summary(model2)
#print(model2, digits=3, corr=FALSE)
#tab_model(model2)

#library(broom.mixed)
#tidy(model2,conf.int=TRUE,exponentiate=TRUE,effects="fixed")
```

**ANOVA:**
```{r , echo=FALSE}
model2aov<- Anova(model2, test="Chisq", type=3)
options(digits=3)
kable(as.data.frame(model2aov))
```


**LS Means - Probability (%):**

```{r , echo=FALSE}
marginal.bin = lsmeans(model2, ~ Treatment)
b <- cld(marginal.bin, alpha   = 0.05, Letters = letters,    
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "sidak")

abc <- data.frame(Treatment = b$Treatment, Probability = b$prob, SE = b$SE, 
                  `Lower CI`= b$asymp.LCL, `Upper CI` = b$asymp.UCL, Group = b$.group)
options(digits=3)
kable(abc)
```


**Differences between treatments:**

```{r , echo=FALSE}

betta = as.data.frame(pairs(marginal.bin, adjust="tukey"))

kable(marginal.bin)
kable(betta)

```


**LS Means - plot:**

```{r, echo=FALSE}
ggplot(abc,
       aes(x     = Treatment,
           y     = Probability,
           color = Treatment,
           label = Group, 
           ymax=1, ymin=0)) +
    geom_point(shape  = 15,
               size   = 4) +
    geom_errorbar(aes(ymin  =  Lower.CI,
                      ymax  =  Upper.CI),
                      width =  0.2,
                      size  =  0.7) +
     theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
    ylab("Probability, %")+
    xlab("Treatment Group")+
   scale_color_manual(values = c("#999999", "#000000", "blue", "red"))+
  theme_bw() +
  theme(legend.position="bottom")

```


**Estimate of risk and risk difference against the NT.**

```{r}
options(digits=5)

### Odds ratio difference: NT - CLOX (Orbenin)
table(datahalf$Treatment, datahalf$IMIatPF)

pr1 = 0.692 
n1 = 43 + 68 #Reference
pr2 = 0.375
n2 = 71 + 46

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

(azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100))



### Odds ratio difference: NT - CEPH (Tomorrow)
table(datahalf$Treatment, datahalf$IMIatPF)

pr1 = 0.692 
n1 = 43 + 68 #Reference
pr2 = 0.404
n2 = 77 + 53

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 


### Odds ratio difference: NT - NC
table(datahalf$Treatment, datahalf$IMIatPF)

pr1 = 0.692 
n1 = 43 + 68 #Reference
pr2 = 0.354
n2 = 96 + 52

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 
```


## Model 3 - Cure:


### Model 3A: Generalized linear mixed model (logistic regression)
 
  - Outcome: Cure.   
  - Fixed effect: Treatment.
  - Covariates: No covariates.
  - Random effects: Farm was included, but removed due to no variation among farms.

**Model:**
```{r , echo=FALSE}
model3 <- glm(Cure ~ Treatment  , 
           family = binomial(link="logit"), 
           #control = glmerControl(optimizer = "bobyqa"),
           data = Desc1)
summary(model3)
#print(model3, digits=3, corr=FALSE)
#tab_model(model3)

```


**ANOVA:**

```{r , echo=FALSE}
model3aov<- Anova(model3,  type=3)

kable(as.data.frame(model3aov))
```


**LS Means - Probability (%):**

```{r , echo=FALSE}
marginal.bin = lsmeans(model3, ~ Treatment)
b <- cld(marginal.bin, alpha   = 0.05, Letters = letters,    
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "sidak")

abcd <- data.frame(Treatment = b$Treatment, Probability = b$prob, SE = b$SE, 
                  `Lower CI`= b$asymp.LCL, `Upper CI` = b$asymp.UCL, Group = b$.group)
kable(abcd)
```


**Differences between treatments:**

```{r , echo=FALSE}

betta = as.data.frame(pairs(marginal.bin, adjust="tukey"))

kable(marginal.bin)
kable(betta)

```


**LS Means - Plot --  plot:** 

```{r, echo=FALSE}
ggplot(abcd,
       aes(x     = Treatment,
           y     = Probability,
           color = Treatment,
           label = Group, 
           ymax=1, ymin=0)) +
    geom_point(shape  = 15,
               size   = 4) +
    geom_errorbar(aes(ymin  =  Lower.CI,
                      ymax  =  Upper.CI),
                      width =  0.2,
                      size  =  0.7) +
     theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
    ylab("Probability, %")+
    xlab("Treatment Group")+
   scale_color_manual(values = c("#999999", "#000000", "blue", "red"))+
  theme_bw() +
  theme(legend.position="bottom")

```


**Estimate of risk and risk difference against NT:**

```{r}
options(digits=5)

### Odds ratio difference: NT - CLOX (Orbenin)
table(datahalf$Treatment, datahalf$Cure)

pr1 = 0.541 
n1 = 51 + 60 #Reference
pr2 = 0.814
n2 = 21 + 92

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

(azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100))



### Odds ratio difference: NT - CEPH (Tomorrow)
table(datahalf$Treatment, datahalf$IMIatPF)

pr1 = 0.541 
n1 = 51 + 60 #Reference
pr2 = 0.854
n2 = 19 + 111

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 

```


**Interpretation:**

  - There is a difference between groups.


## Model 4 - IMI risk:

Statistical model: generalized linear mixed model (logistic regression)

  - Outcome: NewIMI.
  - Fixed effect: Treatment.
  - Covariate: IMI at D0 and Parity tested as confounders, but removed due to non-significance.
  - Random effects: Doe whitin Farms.


**Model:**

```{r , echo=FALSE}
model4 <- glmer(NewIMI ~ Treatment  +  (1 | Farm/DoeID), 
           family = binomial(link="logit"),
           control = glmerControl(optimizer = "bobyqa"),
           data = Desc1)
summary(model4)
#tab_model(model4)
```


**ANOVA:**

```{r , echo=FALSE}
model4aov<- Anova(model4, test="Chisq", type=3)

kable(as.data.frame(model4aov))
```


**LS Means - Probability (%):**

```{r , echo=FALSE}
marginal.bin = lsmeans(model4, ~ Treatment)
b <- cld(marginal.bin, alpha   = 0.05, Letters = letters,    
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "sidak")

ac <- data.frame(Treatment = b$Treatment, Probability = b$prob, SE = b$SE, 
                  `Lower CI`= b$asymp.LCL, `Upper CI` = b$asymp.UCL, Group = b$.group)
kable(ac)
```


**Differences between treatments:**

```{r , echo=FALSE}

betta = as.data.frame(pairs(marginal.bin, adjust="tukey"))

kable(marginal.bin)
kable(betta)

```


**LS Means - Plot:**

```{r, echo=FALSE}
ggplot(ac,
       aes(x     = Treatment,
           y     = Probability,
           color = Treatment,
           label = Group, 
           ymax=1, ymin=0)) +
    geom_point(shape  = 15,
               size   = 4) +
    geom_errorbar(aes(ymin  =  Lower.CI,
                      ymax  =  Upper.CI),
                      width =  0.2,
                      size  =  0.7) +
     theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
    ylab("Probability, %")+
    xlab("Treatment Group")+
   scale_color_manual(values = c("#999999", "#000000", "blue", "red"))+
  theme_bw() +
  theme(legend.position="bottom")

```


**Estimate of risk and risk difference against NT based on logistic regression.**

```{r}
options(digits=5)

### Odds ratio difference: NT - CLOX (Orbenin)
table(datahalf$Treatment, datahalf$NewIMI)

pr1 = 0.32 
n1 = 75 + 36 #Reference
pr2 = 0.25
n2 = 82 + 35

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 


### Odds ratio difference: NT - CEPH (Tomorrow)
table(datahalf$Treatment, datahalf$NewIMI)

pr1 = 0.32 
n1 = 75 + 36 #Reference
pr2 = 0.27
n2 = 90 + 40

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 

### Odds ratio difference: NT - NC 
table(datahalf$Treatment, datahalf$NewIMI)

pr1 = 0.32 
n1 = 75 + 36 #Reference
pr2 = 0.352
n2 = 97 + 52

# Confidence Interval:  
Lower = (pr1-pr2) - 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))
Upper = (pr1-pr2) + 1.96 *(sqrt((pr1*(1-pr1)/n1) + (pr2*(1-pr2)/n2) ))

azz <- data.frame(Risk_Difference =  (pr1-pr2)*100,
                  Lower_CI =  Lower*100, 
                  Upper_CI =  Upper*100)
kable(azz) 

```


# Data analysis - Goat level

## Descriptive statistics:

### By Farm:

```{r, echo=FALSE}

Desc2 <- datagoat %>% subset(select=c(DoeID,	Farm,	Parity, Treatment, DaysDry, MastitisEvent,	Death,	Culling,	DeathCulling)) 

table1::table1(~ DaysDry + Parity + MastitisEvent +	Death +	Culling +	DeathCulling  | Farm,
               data=Desc2)
```


### By Treatment:

```{r, echo=FALSE}

table1::table1(~ DaysDry + MastitisEvent +	Parity + Death +	Culling +	DeathCulling  | Treatment,
               data=Desc2)
```


### Treatment By Farm:

```{r, echo=FALSE}

table1::table1(~  DaysDry + MastitisEvent +	Parity+ Death +	Culling +	DeathCulling 
               | Treatment*Farm,
               data=Desc2)
```


## Risk/Odds of a mastitis event:

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(summarytools)

ctable(x = Desc2$Treatment, 
       y = Desc2$MastitisEvent, useNA = "no", 
       prop = "r")   # Show row proportions
```

**Interpretation:**
  * Since there are no cases of mastitis, statistics cannot be computed.


## Risk/Odds of Culling or Death:

Statistical model: generalized linear mixed model (logistic regression)

  - Outcome: Death or Culling.
  - Fixed effect: Treatment.
  - Covariates: covariates tested and removed if p-value > 0.5 or if there were no changes in the estimates. Parity was tested and removed.
  - Random effects: Farm.

**Model:**

```{r , echo=FALSE}

model6 <- glmer(DeathCulling ~ Treatment +
                  (1 | Farm), 
                family = binomial, control = glmerControl(optimizer = "bobyqa"),
                data = Desc2)

summary(model6)
#print(model6, digits=3, corr=FALSE)
#tab_model(model6

```

**ANOVA:**

```{r , echo=FALSE}
model6aov<- Anova(model6, test="Chisq", type=3)

kable(as.data.frame(model6aov))
```

**LS Means - Probability (%):**

```{r , echo=FALSE}
marginal.bin = lsmeans(model6, ~ Treatment)
b <- cld(marginal.bin, alpha   = 0.05, Letters = letters,    
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "sidak")

abbc <- data.frame(Treatment = b$Treatment, Probability = b$prob, SE = b$SE, 
                  `Lower CI`= b$asymp.LCL, `Upper CI` = b$asymp.UCL, Group = b$.group)
kable(abbc)
```

**Betta and SE:**

```{r , echo=FALSE}

betta = as.data.frame(pairs(marginal.bin, adjust="tukey"))

kable(marginal.bin)
kable(betta)

```

**LS Means - Plot:**

```{r, echo=FALSE}
ggplot(abbc,
       aes(x     = Treatment,
           y     = Probability,
           color = Treatment,
           label = Group,
           ymax=1, ymin=0)) +
    geom_point(shape  = 15,
               size   = 4) +
    geom_errorbar(aes(ymin  =  Lower.CI,
                      ymax  =  Upper.CI),
                      width =  0.2,
                      size  =  0.7) +
     theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
    ylab("Probability, %")+
    xlab("Treatment Group")+
   scale_color_manual(values = c("#999999", "#000000", "blue"))+
  theme_bw()+
  theme(legend.position="bottom")

```


# Models to assess time-to-event:

```{r , echo=FALSE}
# Load the data and remove NAs
datasurv <- datagoat %>% subset(select=c(DoeID,	Farm,	Treatment, DaysDry, SurvMastitis, SurvDeath,	SurvCulling, SurvDeathCulling, DIMatevent)) 
datasurv <- na.omit(datasurv)

```


## Descriptive statistics by treatment:

```{r, echo=FALSE}

table1::table1(~ DaysDry + as.factor(SurvMastitis) +	as.factor(SurvDeath) +	as.factor(SurvCulling) + as.factor(SurvDeathCulling) 
               + DIMatevent  | Treatment,
               data=datasurv)
```


## Model - Survival for Mastitis:

  * Since there are no cases of mastitis, statistics cannot be computed.


## Model - Survival for Culling or Death:

Statistical model: cox regression model (time-to-event) - Culling or Death

  - Outcome: Culling or Death
  - Fixed effect: Treatment.
 

```{r , echo=FALSE}
# Kaplan
fit3 <- survfit(Surv(DIMatevent, SurvDeathCulling) ~ Treatment, data = datasurv)
print(fit3)
ggsurvplot(fit3, data = datasurv) # Survival Plot

## Log-rank
surv_diff <- survdiff(Surv(DIMatevent, SurvDeathCulling) ~ Treatment, data = datasurv)
surv_diff


```

**Survival PLot for Culling or Death.**

```{r , echo=FALSE}
## PLot with more information.
(DeathCulling <- ggsurvplot(fit3, data = datasurv,
 #surv.median.line = "hv", # Add medians survival

 # Change legends: title & labels
 #title = "Cullinging or Death",
 legend = "bottom",

 # Change font size, style and color at the same time
 font.main = c(14, "bold"),
 font.x = c(12, "bold"),
 font.y = c(12, "bold"),
 font.tickslab = c(12, "plain"),
 
 # Add p-value and tervals
 pval = TRUE,

 conf.int = FALSE,
 # Add risk table
 #risk.table = TRUE,
 #tables.height = 0.2,
 #tables.theme = theme_cleantable(),

 # Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
 # or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
 palette = c("#E7B800", "#2E9FDF", "green"),
 ggtheme = theme_bw() # Change ggplot2 theme
))
```


**Cox model - hazard ratio**:

```{r , echo=FALSE}
multi.cox <- coxph(Surv(DIMatevent, SurvDeathCulling) ~ Treatment , 
                   cluster = Farm, data =  datasurv)
summary(multi.cox)

```



# Statistical model for SCC and Milk Yield:

```{r , echo=FALSE}
data_SCCMilk = read_excel("1Final Master Efficacy 8.26.24.xlsx", 
                  sheet = "Goat-level data values", na = ".") 
data_SCCMilk$Treatment  <- as.factor(data_SCCMilk$Treatment)
```


## Descriptive statistics by treatment:

```{r, echo=FALSE}

table1::table1(~ SCCAtDryOff+	SCC1	+SCC2	+SCC3	+SCC4	+DryOffMilkYield	+ `MilkYield_1-30DIM`
               + `MilkYield_31-60DIM` +	`MilkYield_61-90DIM` +	`MilkYield_91-120DIM` 
               | Treatment,
               data=data_SCCMilk)
```


## Somatic cell count (SCC) over Time:

```{r , echo=FALSE}
datascc = read_excel("1Final Master Efficacy 6.24.24.xlsx", 
                  sheet = "SCC", na = ".") 

datascc$Treatment  <- as.factor(datascc$Treatment)
datascc$DoeID <- as.factor(datascc$DoeID)
datascc$QTime <- as.numeric(datascc$Time)
datascc$Time <- as.factor(datascc$Time)
datascc$Farm <- as.factor(datascc$Farm)

```

Statistical model: repeated measures model

  - Outcome: SCC
  - Fixed effect: Treatment and Time.
  - Covariates: covariates tested and removed if p-value > 0.5 or if there were no changes in the estimates.

```{r , echo=FALSE}
model2 <- lme(SCS ~ Treatment * Time,
              random= ~1|DoeID,
              correlation = corCompSymm(form = ~ QTime|DoeID),
              data=datascc, na.action="na.omit")

#summary(model2)
print(model2, digits=3, corr=FALSE)
#tab_model(model2)

#plotNormalHistogram(residuals(model2))

```


**ANOVA:**

```{r , echo=FALSE}
options(digits=3)
anova2 <- Anova(model2, type=3)
kable(as.data.frame(anova2))
```


**LS Means: Interaction between Treatment and Time**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, "Treatment", by="Time", data=datascc)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "sidak")

abc2 <- data.frame(Time = b2$Time, Treatment = b2$Treatment, SCC = b2$lsmean, SEM = b2$SE,
                  Group = b2$.group, lower.CL = b2$lower.CL, upper.CL = b2$upper.CL)
options(digits=5)
kable(abc2)
```

**LS Means: Interaction between Treatment**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, "Treatment", data=datascc)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "Tukey")

pairs(marginal2, adjust="tukey")

kable(b2)

```

**Difference and 95% CI:**

```{r , echo=FALSE}
options(digits=7)

marginal3<- lsmeans(model2, "Treatment", data=datascc)
pairs(marginal3,  adjust="tukey")

# contrast           estimate    SE  df t.ratio p.value
# Control - Orbenin    0.1335 0.210 243   0.635  0.8008

# NT - CLOX
margin.error <- 0.210 * 0.635       # t.ratio*SE
lower.bound <- -0.1335 - margin.error #estimate 
upper.bound <- -0.1335 + margin.error
print(c(lower.bound,upper.bound))


# contrast           estimate    SE  df t.ratio p.value
# Control - Tomorrow   0.1994 0.201 243   0.990  0.5837

# NT - CEPH
margin.error <- 0.201 * 0.990     # t.ratio*SE
lower.bound <- -0.1994 - margin.error #estimate 
upper.bound <- -0.1994 + margin.error
print(c(lower.bound,upper.bound))


```


**LS Means: Interaction between Time**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, "Time", data=datascc)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "Tukey")

pairs(marginal2, adjust="tukey")

kable(b2)

```


**LS Means - Plot 2:**

```{r, echo=FALSE}

(scc_bg <- ggplot(abc2, aes(x=Time , y=SCC, group=Treatment, color=Treatment)) +
    geom_line(size=1.25, position=position_dodge(0.2)) +
    geom_point(size=2, position=position_dodge(0.2))+
    #scale_y_continuous(limits = c(0, 1))+
    #scale_color_manual(values = c("#999999", "#000000"))+
    geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.10, color = NA)+ 
    #ggtitle("Somatic cell count")+
    ylab("SCC")+
    xlab("DIM")+
    scale_x_discrete(labels =c("1-30","31-60","61-90","91-120"))+
    theme_bw()+ 
    theme(legend.position="bottom")+
    theme(plot.title = element_text(size = 14))+ 
    theme(axis.title = element_text(size = 12))  +
    theme(axis.title.x = element_text(size = 12)) +  
    theme(axis.text.x = element_text(size = 12)) +
    theme(axis.text.y = element_text(size = 12)) 
)

```


## Milk Yield over Time:

```{r , echo=FALSE}
datamilk = read_excel("1Final Master Efficacy 8.26.24.xlsx", 
                  sheet = "MilkYield", na = ".") 

datamilk$Treatment  <- as.factor(datamilk$Treatment)
datamilk$DoeID <- as.factor(datamilk$DoeID)
datamilk$QTime <- as.numeric(datamilk$Time)
datamilk$Time <- as.factor(datamilk$Time)
datamilk$Farm <- as.factor(datamilk$Farm)

```


Statistical model: repeated measures model

  - Outcome: Milk Yield
  - Fixed effect: Treatment and Time.
  - Covariates: covariates tested and removed if p-value > 0.5 or if there were no changes in the estimates.

```{r , echo=FALSE}
model2 <- lme(MilkYield ~ Treatment * Time,
              random= ~1|DoeID/Farm,
              correlation = corCompSymm(form = ~ QTime|DoeID),
              data=datamilk, na.action="na.omit")

#summary(model2)
print(model2, digits=3, corr=FALSE)
#tab_model(model2)

#plotNormalHistogram(residuals(model2))
```


**ANOVA:**

```{r , echo=FALSE}
anova2 <- Anova(model2, type=3)
kable(as.data.frame(anova2))
```


**LS Means:**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, "Treatment", by="Time", data=datamilk)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "sidak")

as.data.frame(pairs(marginal2, adjust="tukey"))

abc2 <- data.frame(Time = b2$Time, Treatment = b2$Treatment, Yield = b2$lsmean, SEM = b2$SE,
                  Group = b2$.group, lower.CL = b2$lower.CL, upper.CL = b2$upper.CL)
options(digits=4)
kable(abc2)
```


**LS Means - Plot 1:**

```{r, echo=FALSE}

ggplot(abc2,
       aes(x     = Time,
           y     = Yield,
           color = Treatment,
           label = Group)) +
  geom_line() +
    geom_point(shape  = 15,
               size   = 4) +
    geom_errorbar(aes(ymin  =  lower.CL,
                      ymax  =  upper.CL),
                      width =  0.2,
                      size  =  0.7) +
     theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
    theme(legend.position="bottom")+
    ylab("Milk Yield")+
    xlab("Time")


```

**LS Means - Plot 2:**

```{r, echo=FALSE}

(scc_bg <- ggplot(abc2, aes(x=Time , y=Yield, group=Treatment, color=Treatment)) +
    geom_line(size=1.25, position=position_dodge(0.2)) +
    geom_point(size=2, position=position_dodge(0.2))+
    #scale_y_continuous(limits = c(0, 1))+
    #scale_color_manual(values = c("#999999", "#000000"))+
    geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.10, color = NA)+ 
    #ggtitle("Somatic cell count")+
    ylab("MilkYield (lbs/day)")+
    xlab("DIM")+
    scale_x_discrete(labels =c("1-30 DIM","31-60 DIM"," 61-90 DIM","91-120 DIM"))+
    theme_bw()+ 
    theme(legend.position="bottom")+
    theme(plot.title = element_text(size = 14))+ 
    theme(axis.title = element_text(size = 12))  +
    theme(axis.title.x = element_text(size = 12)) +  
    theme(axis.text.x = element_text(size = 12)) +
    theme(axis.text.y = element_text(size = 12)) 
)

```

**LS Means - Treatment difference:**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, "Treatment", data=datascc)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "Tukey")

kable(b2)

```


**LS Means - Treatment difference:**

```{r}

pairs(marginal2, adjust="tukey")


# CEPH - NT
margin.error <- -0.480 * 0.633
lower.bound <- -0.3037 - margin.error
upper.bound <- -0.3037 + margin.error
print(c(lower.bound,upper.bound))

# CLOX - NT
margin.error <- -0.378 * 0.666
lower.bound <- -0.2515 - margin.error
upper.bound <- -0.2515 + margin.error
print(c(lower.bound,upper.bound))
```



Statistical model: repeated measures model

  - Outcome: MilkYieldKg
  - Fixed effect: Treatment and Time.
  - Covariates: covariates tested and removed if p-value > 0.5 or if there were no changes in the estimates.

```{r , echo=FALSE}
model2 <- lme(MilkYieldKg ~ Treatment * Time,
              random= ~1|DoeID,
              correlation = corCompSymm(form = ~ QTime|DoeID),
              data=datamilk, na.action="na.omit")

#summary(model2)
print(model2, digits=3, corr=FALSE)
#tab_model(model2)

#plotNormalHistogram(residuals(model2))
```


**ANOVA:**

```{r , echo=FALSE}
anova2 <- Anova(model2, type=3)
kable(as.data.frame(anova2))
```


**LS Means:**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, "Treatment", by="Time", data=datamilk)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "sidak")

as.data.frame(pairs(marginal2, adjust="tukey"))

abc2 <- data.frame(Time = b2$Time, Treatment = b2$Treatment, Yield = b2$lsmean, SEM = b2$SE,
                  Group = b2$.group, lower.CL = b2$lower.CL, upper.CL = b2$upper.CL)
options(digits=4)
kable(abc2)
```


**LS Means - Plot 1:**

```{r, echo=FALSE}

ggplot(abc2,
       aes(x     = Time,
           y     = Yield,
           color = Treatment,
           label = Group)) +
  geom_line() +
    geom_point(shape  = 15,
               size   = 4) +
    geom_errorbar(aes(ymin  =  lower.CL,
                      ymax  =  upper.CL),
                      linewidth =  0.2,
                      size  =  0.7) +
     theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
    theme(legend.position="bottom")+
    ylab("Milk Yield")+
    xlab("Time")


```

**LS Means - Plot 2:**

```{r, echo=FALSE}

(scc_bg <- ggplot(abc2, aes(x=Time , y=Yield, group=Treatment, color=Treatment)) +
    geom_line(size=1.25, position=position_dodge(0.2)) +
    geom_point(size=2, position=position_dodge(0.2))+
    #scale_y_continuous(limits = c(0, 1))+
    #scale_color_manual(values = c("#999999", "#000000"))+
    geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.10, color = NA)+ 
    #ggtitle("Somatic cell count")+
    ylab("MilkYield (kgs/day)")+
    xlab("DIM")+
    scale_x_discrete(labels =c("1-30 DIM","31-60 DIM"," 61-90 DIM","91-120 DIM"))+
    theme_bw()+ 
    theme(legend.position="bottom")+
    theme(plot.title = element_text(size = 14))+ 
    theme(axis.title = element_text(size = 12))  +
    theme(axis.title.x = element_text(size = 12)) +  
    theme(axis.text.x = element_text(size = 12)) +
    theme(axis.text.y = element_text(size = 12)) 
)

```

**LS Means - Treatment difference:**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, "Treatment", data=datamilk)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "Tukey")

kable(b2)

```

**LS Means - Treatment difference:**

```{r , echo=FALSE}
marginal2<- lsmeans(model2, Time, data=datamilk)
#kable(marginal2)

b2 <- cld(marginal2, alpha   = 0.05, Letters = letters,    ### Use lower-case letters for .group
    type    = "response", ### Report emmeans in orginal scale
    adjust =  "Tukey")

kable(b2)

```


